### 2022-05-29 LaTeX/report.tex
Partial_correctness and report updates

\theoremstyle{definition}
\newtheorem{isabelle}{Isabelle}

@@

\isa{
    {\color{isa_blue}record} \generic{v} env =\\
    $~~~\mathcal{S}$ :: "\generic{v} $\Rightarrow$ \generic{v} set"\\
    $~~~$explored :: "\generic{v} set"\\
    $~~~$visited :: "\generic{v} set"\\
    $~~~$sccs :: "\generic{v} set set"\\
    $~~~$stack :: "\generic{v} list"
}

@@

\isa{
    {\locale} graph =\\
    $~~~$\fixes{} vertices :: "\generic{v} set" \and{} successors :: "\generic{v} $\Rightarrow$ \generic{v} set"\\
    $~~~$\assumes{} vfin: "finite vertices"\\
    $~~~$\and{} sclosed: "$\forall$ \green{x} $\in$ vertices. successors \green{x} $\subseteq$ vertices"
}

@@

\isa{
    \abbreviation{} edge \where\\
    $~~~$"edge \green{x y} $\equiv$ \green{y} $\in$ successors \green{x}"
}

@@

\noindent
Regarding the reachability binary relation, a choice has to be made since there are several ways to define it. In particular, there are two possible keywords, {\color{isa_blue}\texttt{inductive}} and {\color{isa_blue}\texttt{fun}}, respectively for an inductive or recursive definition. If both definitions are valid, the inductive one is kept for the following reasons. Although a recursive definition has to be based on some underlying inductive definition -- which is simply not available for graphs -- it expresses both the positive and negative information\footnote{In this case, the positive information designates the fact of being reachable and the negative information designates the fact of not being reachable.} whereas the inductive one only expresses the positive information directly. Therefore, with an inductive definition, the negative information has to be proved. One would be right to argue that it would be more convenient to be able to tell without proving it that two nodes are not reachable from each other, but this does not interest us for the following. Another important point is that there is no datatype for a recursive definition, especially in this case with the transitive closure of the $\Rightarrow^*$ relation. Thus, the choice of the inductive definition is not a choice of simplicity but of necessity. Isabelle provides a construction for inductive predicate definitions, which is appropriate here because the two clauses represent the reflexive case and the extension of reachability by prepending an edge. This will be particularly useful in the proofs.

\isa{
    \inductive{} reachable \where\\
    $~~~$reachable\_refl[iff]: "reachable \green{x x}"\\
    $~\mid$ reachable\_succ[elim]: "$\ldb$edge \green{x y}; reachable \green{y z}$\rdb \Longrightarrow$ reachable \green{x z}"\\
}

@@

\isa{
\lm{} succ\_reachable:\\
$~~~~~$\assumes{} "reachable {\color{isa_dark_blue}x y}" \and{} "edge {\color{isa_dark_blue}y z}"\\
$~~~~~$\shows "reachable {\color{isa_dark_blue}x z}"\\
$~~~~~${\color{isa_blue}using} assms {\color{isa_blue}by} induct auto\\
}
Mathematical writing: $\forall x, \forall y, \forall z, (x \Rightarrow^* y \wedge y \Rightarrow z) \Longrightarrow x \Rightarrow^* z$

@@

\isa{
\lm{} reachable\_trans:\\
$~~~~~$\assumes{} y: "reachable \blue{x y}" \and{} z: "reachable \blue{y z}"\\
$~~~~~$\shows{} "reachable \blue{x z}"\\
$~~~~~${\color{isa_blue}using} assms {\color{isa_blue}by} induct auto\\
}
Mathematical writing: $\forall x, \forall y, \forall z, (x \Rightarrow^* y \wedge y \Rightarrow^* z) \Longrightarrow x \Rightarrow^* z$\\

@@

\isa{
{\color{isa_blue}definition} is\_subscc \where\\
$~~~~~$"is\_subscc S $\equiv \forall$ x $\in$ S. $\forall$ y $\in$ S. reachable x y"\\
}
Mathematical writing: A set $S$ is a SCC if $\forall x \in S, \forall y \in S, x \Rightarrow^* y$

\isa{
{\color{isa_blue}definition} is\_scc \where\\
$~~~~~$"is\_scc S $\equiv$ S $\neq$ \{\} $\wedge$ is\_subscc S\\
$~~~~~\wedge$ ($\forall$ S'. S $\subseteq$ S' $\wedge$ is\_subscc S' $\longrightarrow$ S' = S)"\\
}
Mathematical writing: A non-empty SCC $S$ is maximal if for all SCC $S'$, $S \subseteq S' \Longrightarrow S'=S$\\

@@

The idea is to later use the following property: if $x \preceq y~\text{in}~xs$, then $y \Rightarrow^* x$.\\
It is defined in Isabelle as follows:\\
\isa{
    {\color{isa_blue}definition} precedes ("\_ $\preceq$ \_ in \_" [100,100,100] 39) \where\\
$~~~~~$"{\color{isa_dark_green}x} $\preceq$ {\color{isa_dark_green}y} in {\color{isa_dark_green}xs} $\equiv$ $\exists${\color{isa_dark_green}h} {\color{isa_dark_green}r}. {\color{isa_dark_green}xs} = {\color{isa_dark_green}h} @ ({\color{isa_dark_green}x} \# {\color{isa_dark_green}r}) $\wedge$ {\color{isa_dark_green}y} $\in$ set ({\color{isa_dark_green}x} \# {\color{isa_dark_green}r})"\\
}

@@

\isa{
    {\color{isa_blue}function} unite :: "\generic{v} $\Rightarrow$ \generic{v} $\Rightarrow$ \generic{v} env $\Rightarrow$ \generic{v} env" \where{}\\
    "\blue{unite} \green{v w e} =\\
    $~~~$({\color{isa_blue}if} ($\mathcal{S}$ \green{e v} = $\mathcal{S}$ \green{e w}) \bblue{then} \green{e}\\
    $~~~$\bblue{else let} \green{r} = hd(stack \green{e});\\
    $~~~~~~~~~~~~$\green{r'}= hd(tl(stack \green{e}));\\
    $~~~~~~~~~~~~$\green{joined} = $\mathcal{S}$ \green{e r} $\cup$ $\mathcal{S}$ \green{e r};\\
    $~~~~~~~~~~~~$\green{e'}= \green{e}\env{\\
    $~~~~~~~~~~~~~~~$stack := tl(stack \green{e}),\\
    $~~~~~~~~~~~~~~~\mathcal{S}$ := ($\lambda$ n. \bblue{if} n $\in$ \green{joined} \bblue{then} \green{joined} \bblue{else} $\mathcal{S}$ \green{e n})\\
    $~~~~~~~~~~~~$}\\
    $~~~$\bblue{in} \blue{unite} \green{v w e'})"\\
    $~~~$\bblue{by} pat\_completeness auto
}

@@

\isa{
    \bblue{definition} unite :: "\generic{v} $\Rightarrow$ \generic{v} $\Rightarrow$ \generic{v} env $\Rightarrow$ \generic{v} env" \where{}\\
    $~~~$"\blue{unite} \green{v w e} $\equiv$\\
    $~~~~~~$\bblue{let} \green{pfx} = takeWhile ($\lambda$\green{x}. \green{w} $\notin$ $\mathcal{S}$ \green{e x}) (stack \green{e});\\
    $~~~~~~~~~~$\green{sfx} = dropWhile ($\lambda$\green{x}. \green{w} $\notin$ $\mathcal{S}$ \green{e x}) (stack \green{e});\\
    $~~~~~~~~~~$\green{cc} = $\bigcup$ \{$\mathcal{S}$ \green{e x} $|$ \green{x}. \green{x} $\in$ set \green{pfx} $\cup$ \{hd \green{sfx}\}\}\\
    $~~~~~~$\bblue{in}  \green{e}\env{$\mathcal{S}$ := $\lambda$\green{x}. \bblue{if} \green{x} $\in$ \green{cc} \bblue{then} \green{cc} \bblue{else} $\mathcal{S}$ \green{e x}, stack := \green{sfx}}"
}

@@

\isa{
    {\color{isa_blue}{function}} dfs :: "\generic{v} $\Rightarrow$ \generic{v} env $\Rightarrow$ \generic{v} env" \and{}\\
    $~~~~~~~~~~$dfss:: "\generic{v} $\Rightarrow$ \generic{v} set $\Rightarrow$ \generic{v} env $\Rightarrow$ \generic{v} env" \where\\
    "{\color{isa_dark_blue}dfs} {\color{isa_dark_green}v e} =\\
    $~~~~~$({\color{isa_blue}{let}} {\color{isa_dark_green}e1} = {\color{isa_dark_green}e}\env{visited := visited e $\cup$ \{{\color{isa_dark_green}v}\}, stack := ({\color{isa_dark_green}v} \# stack {\color{isa_dark_green}e})};\\
    $~~~~~~~~~~${\color{isa_dark_green}e'} = {\color{isa_dark_blue}dfss} {\color{isa_dark_green}v} (successors {\color{isa_dark_green}v}) {\color{isa_dark_green}e1}\\
    $~~~~~${\color{isa_blue}{in if}} {\color{isa_dark_green}v} = hd(stack {\color{isa_dark_green}e'})\\
    $~~~~~~~~~~${\color{isa_blue}{then}} {\color{isa_dark_green}e'}\env{sccs:=sccs {\color{isa_dark_green}e'} $\cup$ {$\mathcal{S}$ {\color{isa_dark_green}e' v}}, explored:=explored {\color{isa_dark_green}e'} $\cup$ ($\mathcal{S}$ {\color{isa_dark_green}e' v}), stack:=tl(stack {\color{isa_dark_green}e'})}\\
    $~~~~~~~~~~${\color{isa_blue}else} {\color{isa_dark_green}e'})"\\
    | "{\color{isa_dark_blue}dfss} {\color{isa_dark_green}v vs e} =\\
    $~~~~~$({\color{isa_blue}if} {\color{isa_dark_green}vs} = \{\} {\color{isa_blue}then} {\color{isa_dark_green}e}\\
    $~~~~~${\color{isa_blue}else} ({\color{isa_blue}let} {\color{isa_dark_green}w} = SOME {\color{isa_dark_green}x}. {\color{isa_dark_green}x} $\in$ {\color{isa_dark_green}vs}\\
    $~~~~~~~~~~${\color{isa_blue}in} ({\color{isa_blue}let} {\color{isa_dark_green}e'} = ({\color{isa_blue}if} {\color{isa_dark_green}w} $\in$ explored {\color{isa_dark_green}e} {\color{isa_blue}then} {\color{isa_dark_green}e}\\
    $~~~~~~~~~~~~~~~~~~~${\color{isa_blue}else if} {\color{isa_dark_green}w} $\notin$ visited {\color{isa_dark_green}e} {\color{isa_blue}then} {\color{isa_dark_blue}dfs} {\color{isa_dark_green}w e}\\
    $~~~~~~~~~~~~~~~~~~~${\color{isa_blue}else} unite {\color{isa_dark_green}v w e})\\
    $~~~~~~~~~~~~~~${\color{isa_blue}in} {\color{isa_dark_blue}dfss} {\color{isa_dark_green}v} ({\color{isa_dark_green}v} - \{{\color{isa_dark_green}w}\}) {\color{isa_dark_green}e'})))"\\
    $~~~${\color{isa_blue}by} pat\_completeness (force\bblue{+})\\
}

@@

\isa{
    \bblue{definition} wf\_env \where{}\\
    $~~~$"\blue{wf\_env} \green{e} $\equiv$\\
    $~~~~~~$distinct (stack \green{e})\\
    $~~~~\wedge$ set (stack \green{e}) $\subseteq$ visited \green{e}\\
    $~~~~\wedge$ explored \green{e} $\subseteq$ visited \green{e}\\
    $~~~~\wedge$ explored \green{e} $\cap$ set (stack \green{e}) = \{\}\\
    $~~~~\wedge$ ($\forall$ \green{v} \green{w}. \green{w} $\in$ $\mathcal{S}$ \green{e} \green{v} $\longleftrightarrow$ ($\mathcal{S}$ \green{e} \green{v} = $\mathcal{S}$ \green{e} \green{w}))\\
    $~~~~\wedge$ ($\forall$\green{v} $\in$ set (stack \green{e}).$\forall$ \green{w} $\in$ set (stack \green{e}).\green{v} $\neq$ \green{w} $\longrightarrow$ $\mathcal{S}$ \green{e} \green{v} $\cap$ $\mathcal{S}$ \green{e} \green{w} = \{\})\\
    $~~~~\wedge$ ($\forall$ \green{v}. \green{v} $\notin$ visited \green{e} $\longrightarrow$ $\mathcal{S}$ \green{e} \green{v} = \{\green{v}\})\\
    $~~~~\wedge$ $\bigcup$ \{$\mathcal{S}$ \green{e} \green{v} $|$ \green{v}. \green{v} $\in$ set (stack \green{e})\} = visited \green{e} - explored \green{e}\\
    $~~~~\wedge$ ($\forall$ \green{x} \green{y}. \green{x} $\preceq$ \green{y} in stack \green{e} $\longrightarrow$ reachable \green{y} \green{x})\\
    $~~~~\wedge$ ($\forall$ \green{x}. is\_subscc ($\mathcal{S}$ \green{e} \green{x}))\\
    $~~~~\wedge$ ($\forall$ \green{x} $\in$ explored \green{e}. $\forall$ \green{y}. reachable \green{x} \green{y} $\longrightarrow$ \green{y} $\in$ explored \green{e})\\
    $~~~~\wedge$ ($\forall$ \green{S} $\in$ sccs \green{e}. is\_scc \green{S})"
}

@@

\isa{
\bblue{definition} sub\_env \where\\
"\blue{sub\_env} \green{e e'} $\equiv$\\
$~~~~$visited \green{e} $\subseteq$ visited \green{e'}\\
$~~$$\wedge$ explored \green{e} $\subseteq$ explored \green{e'}\\
$~~$$\wedge$ ($\forall$ \green{v}. $\mathcal{S}$ \green{e v} $\subseteq$ $\mathcal{S}$ \green{e' v})\\
$~~$$\wedge$ ($\bigcup$\{$\mathcal{S}$ \green{e v} $|$ \green{v}. \green{v} $\in$ set (stack \green{e})\}) $\subseteq$ ($\bigcup$\{$\mathcal{S}$ \green{e' v} $|$ \green{v}. \green{v} $\in$ set (stack \green{e'})\})"
}

@@

\isa{
\bblue{definition} post\_dfs \where\\ 
"post\_dfs \green{v prev\_e e} $\equiv$\\
$~~~~$wf\_env \green{e}\\
$~~\wedge$ ($\forall$ \green{x}. reachable \green{v} \green{x} $\longrightarrow$ \green{x} $\in$ visited e)\\
$~~\wedge$ sub\_env \green{prev\_e} \green{e}\\
$~~\wedge$ ($\forall$ \green{n} $\in$ set (stack \green{e}). reachable \green{n} \green{v})\\
$~~\wedge$ ($\exists$ \green{ns}. stack \green{prev\_e} = \green{ns} @ (stack \green{e}))\\
$~~\wedge$ ($\forall$ \green{m n}. $\stackprec{\green{m}}{\green{n}}{\green{prec\_e}}$ $\longrightarrow$ \\ $~~~~~~$ ($\forall$ \green{u} $\in$ $\mathcal{S}$ \green{prev\_e m}. reachable \green{u} \green{v} $\wedge$ reachable \green{v} \green{n} $\longrightarrow$ $\mathcal{S}$ \green{e m} = $\mathcal{S}$ \green{e n}))\\
$~~\wedge$ ((\green{v} $\in$ explored \green{e} $\wedge$ stack \green{e} = stack \green{prev\_e}) $\lor$ \\
$~~~~~~$ (\green{v} $\in$ $\mathcal{S}$ \green{e} (hd (stack \green{e}))) $\wedge$ \\
$~~~~~~$ ($\exists$ \green{n} $\in$ set (stack \green{prev\_e}). $\mathcal{S}$ \green{e v} = $\mathcal{S}$ \green{e n}))"
}

@@

\isa{
\bblue{definition} pre\_dfss \where{}\\
"pre\_dfss \green{v vs e} $\equiv$\\
$~~~~$wf\_env \green{e}\\
$~~\land$ \green{v} $\in$ visited \green{e}\\
$~~\land$ \green{vs} $\subseteq$ successors \green{v}\\
$~~\land$ ($\forall$ \green{n} $\in$ set (stack \green{e}). reachable \green{n v})\\
$~~\land$ (stack \green{e} $\neq$ [])\\
$~~\land$ ($\forall$ \green{n}$\in$set(stack \green{e}).reachable \green{v n} $\longrightarrow$ \green{v}$\in$$\mathcal{S}$ \green{e n} $\lor$ ($\exists$ \green{m} $\in$ \green{vs}.reachable \green{m n}))\\
$~~\land$ (\green{v} $\in$ $\mathcal{S}$ \green{e} (hd (stack \green{e})))"
}

@@

\isa{
\bblue{definition} post\_dfss \where\\
"post\_dfss \green{v vs prev\_e e} $\equiv$\\
$~~~~$wf\_env \green{e}\\
$~~\land$ ($\forall$ \green{w} $\in$ \green{vs}. $\forall$ \green{x}. reachable \green{w x} $\longrightarrow$ \green{x} $\in$ visited \green{e})\\
$~~\land$ sub\_env \green{prev\_e e}\\
$~~\land$ ($\forall$ \green{n} $\in$ set (stack \green{e}). reachable \green{n v})\\
$~~\land$ (stack \green{e} $\neq$ [])\\
$~~\land$ ($\forall$ \green{n} $\in$ set (stack \green{e}). reachable \green{v n} $\longrightarrow$ \green{v} $\in$ $\mathcal{S}$ \green{e n})\\
$~~\land$ ($\exists$ \green{ns}. stack \green{prev\_e} = \green{ns} @ (stack \green{e}))\\
$~~\land$ (\green{v} $\in$ $\mathcal{S}$ \green{e} (hd (stack \green{e})))"
}

\subsection{Formal proof}
\subsubsection{\texttt{pre\_dfs} implies \texttt{pre\_dfss}}

>>>>

\newcommand{\isabelle}[1]{
    \noindent\begin{minipage}{1.1\linewidth}
        \isa{
            #1
        }
    \end{minipage}
}

@@

\isabelle{
    {\color{isa_blue}record} \generic{v} env =\\
    $~~~\mathcal{S}$ :: "\generic{v} $\Rightarrow$ \generic{v} set"\\
    $~~~$explored :: "\generic{v} set"\\
    $~~~$visited :: "\generic{v} set"\\
    $~~~$sccs :: "\generic{v} set set"\\
    $~~~$stack :: "\generic{v} list"
}

@@

\isabelle{
    {\locale} graph =\\
    $~~~$\fixes{} vertices :: "\generic{v} set" \and{} successors :: "\generic{v} $\Rightarrow$ \generic{v} set"\\
    $~~~$\assumes{} vfin: "finite vertices"\\
    $~~~$\and{} sclosed: "$\forall$ \green{x} $\in$ vertices. successors \green{x} $\subseteq$ vertices"
}

@@

\isabelle{
    \abbreviation{} edge \where\\
    $~~~$"edge \green{x y} $\equiv$ \green{y} $\in$ successors \green{x}"
}

@@

\noindent

Isabelle allows the definition of recursive functions. Such a definition must guarantee that any recursive call takes an argument that decreases (according to some measure function) with respect to the original argument. This is typically the case for functions that follow the definition of an algebraic type (integers, lists, trees, etc.). In addition, Isabelle allows inductive definitions of relations where two values are related if it is possible to justify this relation by applying a finite number of times the definition clauses of the relation. In our case, this applies to the definition of reachability where two nodes are connected if they are identical or if the first node has a successor from which the second node is already known to be reachable.

Although a recursive definition has to be based on some underlying inductive definition -- which is simply not available for graphs -- it expresses both the positive and negative information\footnote{In this case, the positive information designates the fact of being reachable and the negative information designates the fact of not being reachable.} whereas the inductive one only expresses the positive information directly. Therefore, with an inductive definition, the negative information has to be proved. One would be right to argue that it would be more convenient to be able to tell without proving it that two nodes are not reachable from each other, but this does not interest us for the following and this is actually more difficult than proving the reachability. Another important point is that there is no datatype for a recursive definition, especially in this case with the transitive closure of the $\Rightarrow^*$ relation. Thus, the inductive definition is not a choice but is necessary. Isabelle provides a construction for inductive predicate definitions, which is appropriate here because the two clauses represent the reflexive case and the extension of reachability by prepending an edge. This will be particularly useful in the proofs.

\isabelle{
    \inductive{} reachable \where\\
    $~~~$reachable\_refl[iff]: "reachable \green{x x}"\\
    $~\mid$ reachable\_succ[elim]: "$\ldb$edge \green{x y}; reachable \green{y z}$\rdb \Longrightarrow$ reachable \green{x z}"\\
}

@@

\isabelle{
\lm{} succ\_reachable:\\
$~~~~~$\assumes{} "reachable {\color{isa_dark_blue}x y}" \and{} "edge {\color{isa_dark_blue}y z}"\\
$~~~~~$\shows "reachable {\color{isa_dark_blue}x z}"\\
$~~~~~${\color{isa_blue}using} assms {\color{isa_blue}by} induct auto\\
}
Mathematical writing: $\forall x, \forall y, \forall z, (x \Rightarrow^* y \wedge y \Rightarrow z) \Longrightarrow x \Rightarrow^* z$

@@

\isabelle{
\lm{} reachable\_trans:\\
$~~~~~$\assumes{} y: "reachable \blue{x y}" \and{} z: "reachable \blue{y z}"\\
$~~~~~$\shows{} "reachable \blue{x z}"\\
$~~~~~${\color{isa_blue}using} assms {\color{isa_blue}by} induct auto\\
}
Mathematical writing: $\forall x, \forall y, \forall z, (x \Rightarrow^* y \wedge y \Rightarrow^* z) \Longrightarrow x \Rightarrow^* z$\\

@@

\isabelle{
{\color{isa_blue}definition} is\_subscc \where\\
$~~~~~$"is\_subscc S $\equiv \forall$ x $\in$ S. $\forall$ y $\in$ S. reachable x y"\\
}
Mathematical writing: A set $S$ is a SCC if $\forall x \in S, \forall y \in S, x \Rightarrow^* y$

\isabelle{
{\color{isa_blue}definition} is\_scc \where\\
$~~~~~$"is\_scc S $\equiv$ S $\neq$ \{\} $\wedge$ is\_subscc S\\
$~~~~~\wedge$ ($\forall$ S'. S $\subseteq$ S' $\wedge$ is\_subscc S' $\longrightarrow$ S' = S)"\\
}
Mathematical writing: A non-empty SCC $S$ is maximal if for all SCC $S'$, $S \subseteq S' \Longrightarrow S'=S$\\

@@

The idea is to later use the following property: if $x \preceq y~\text{in}~xs$, then $y \Rightarrow^* x$.\\
It is defined in Isabelle as follows:\\

\isabelle{
    {\color{isa_blue}definition} precedes ("\_ $\preceq$ \_ in \_" [100,100,100] 39) \where\\
$~~~~~$"{\color{isa_dark_green}x} $\preceq$ {\color{isa_dark_green}y} in {\color{isa_dark_green}xs} $\equiv$ $\exists${\color{isa_dark_green}h} {\color{isa_dark_green}r}. {\color{isa_dark_green}xs} = {\color{isa_dark_green}h} @ ({\color{isa_dark_green}x} \# {\color{isa_dark_green}r}) $\wedge$ {\color{isa_dark_green}y} $\in$ set ({\color{isa_dark_green}x} \# {\color{isa_dark_green}r})"\\
}

@@

\isabelle{
    {\color{isa_blue}function} unite :: "\generic{v} $\Rightarrow$ \generic{v} $\Rightarrow$ \generic{v} env $\Rightarrow$ \generic{v} env" \where{}\\
    "\blue{unite} \green{v w e} =\\
    $~~~$({\color{isa_blue}if} ($\mathcal{S}$ \green{e v} = $\mathcal{S}$ \green{e w}) \bblue{then} \green{e}\\
    $~~~$\bblue{else let} \green{r} = hd(stack \green{e});\\
    $~~~~~~~~~~~~$\green{r'}= hd(tl(stack \green{e}));\\
    $~~~~~~~~~~~~$\green{joined} = $\mathcal{S}$ \green{e r} $\cup$ $\mathcal{S}$ \green{e r};\\
    $~~~~~~~~~~~~$\green{e'}= \green{e}\env{\\
    $~~~~~~~~~~~~~~~$stack := tl(stack \green{e}),\\
    $~~~~~~~~~~~~~~~\mathcal{S}$ := ($\lambda$ n. \bblue{if} n $\in$ \green{joined} \bblue{then} \green{joined} \bblue{else} $\mathcal{S}$ \green{e n})\\
    $~~~~~~~~~~~~$}\\
    $~~~$\bblue{in} \blue{unite} \green{v w e'})"\\
    $~~~$\bblue{by} pat\_completeness auto
}

@@

\isabelle{
    \bblue{definition} unite :: "\generic{v} $\Rightarrow$ \generic{v} $\Rightarrow$ \generic{v} env $\Rightarrow$ \generic{v} env" \where{}\\
    $~~~$"\blue{unite} \green{v w e} $\equiv$\\
    $~~~~~~$\bblue{let} \green{pfx} = takeWhile ($\lambda$\green{x}. \green{w} $\notin$ $\mathcal{S}$ \green{e x}) (stack \green{e});\\
    $~~~~~~~~~~$\green{sfx} = dropWhile ($\lambda$\green{x}. \green{w} $\notin$ $\mathcal{S}$ \green{e x}) (stack \green{e});\\
    $~~~~~~~~~~$\green{cc} = $\bigcup$ \{$\mathcal{S}$ \green{e x} $|$ \green{x}. \green{x} $\in$ set \green{pfx} $\cup$ \{hd \green{sfx}\}\}\\
    $~~~~~~$\bblue{in}  \green{e}\env{$\mathcal{S}$ := $\lambda$\green{x}. \bblue{if} \green{x} $\in$ \green{cc} \bblue{then} \green{cc} \bblue{else} $\mathcal{S}$ \green{e x}, stack := \green{sfx}}"
}

@@

\isabelle{
    {\color{isa_blue}{function}} dfs :: "\generic{v} $\Rightarrow$ \generic{v} env $\Rightarrow$ \generic{v} env" \and{}\\
    $~~~~~~~~~~$dfss:: "\generic{v} $\Rightarrow$ \generic{v} set $\Rightarrow$ \generic{v} env $\Rightarrow$ \generic{v} env" \where\\
    "{\color{isa_dark_blue}dfs} {\color{isa_dark_green}v e} =\\
    $~~~~~$({\color{isa_blue}{let}} {\color{isa_dark_green}e1} = {\color{isa_dark_green}e}\env{visited := visited e $\cup$ \{{\color{isa_dark_green}v}\}, stack := ({\color{isa_dark_green}v} \# stack {\color{isa_dark_green}e})};\\
    $~~~~~~~~~~${\color{isa_dark_green}e'} = {\color{isa_dark_blue}dfss} {\color{isa_dark_green}v} (successors {\color{isa_dark_green}v}) {\color{isa_dark_green}e1}\\
    $~~~~~${\color{isa_blue}{in if}} {\color{isa_dark_green}v} = hd(stack {\color{isa_dark_green}e'})\\
    $~~~~~~~~~~${\color{isa_blue}{then}} {\color{isa_dark_green}e'}\env{sccs:=sccs {\color{isa_dark_green}e'} $\cup$ {$\mathcal{S}$ {\color{isa_dark_green}e' v}}, explored:=explored {\color{isa_dark_green}e'} $\cup$ ($\mathcal{S}$ {\color{isa_dark_green}e' v}), stack:=tl(stack {\color{isa_dark_green}e'})}\\
    $~~~~~~~~~~${\color{isa_blue}else} {\color{isa_dark_green}e'})"\\
    | "{\color{isa_dark_blue}dfss} {\color{isa_dark_green}v vs e} =\\
    $~~~~~$({\color{isa_blue}if} {\color{isa_dark_green}vs} = \{\} {\color{isa_blue}then} {\color{isa_dark_green}e}\\
    $~~~~~${\color{isa_blue}else} ({\color{isa_blue}let} {\color{isa_dark_green}w} = SOME {\color{isa_dark_green}x}. {\color{isa_dark_green}x} $\in$ {\color{isa_dark_green}vs}\\
    $~~~~~~~~~~${\color{isa_blue}in} ({\color{isa_blue}let} {\color{isa_dark_green}e'} = ({\color{isa_blue}if} {\color{isa_dark_green}w} $\in$ explored {\color{isa_dark_green}e} {\color{isa_blue}then} {\color{isa_dark_green}e}\\
    $~~~~~~~~~~~~~~~~~~~${\color{isa_blue}else if} {\color{isa_dark_green}w} $\notin$ visited {\color{isa_dark_green}e} {\color{isa_blue}then} {\color{isa_dark_blue}dfs} {\color{isa_dark_green}w e}\\
    $~~~~~~~~~~~~~~~~~~~${\color{isa_blue}else} unite {\color{isa_dark_green}v w e})\\
    $~~~~~~~~~~~~~~${\color{isa_blue}in} {\color{isa_dark_blue}dfss} {\color{isa_dark_green}v} ({\color{isa_dark_green}v} - \{{\color{isa_dark_green}w}\}) {\color{isa_dark_green}e'})))"\\
    $~~~${\color{isa_blue}by} pat\_completeness (force\bblue{+})\\
}

@@

\isabelle{
    \bblue{definition} wf\_env \where{}\\
    $~~~$"\blue{wf\_env} \green{e} $\equiv$\\
    $~~~~~~$distinct (stack \green{e})\\
    $~~~~\wedge$ set (stack \green{e}) $\subseteq$ visited \green{e}\\
    $~~~~\wedge$ explored \green{e} $\subseteq$ visited \green{e}\\
    $~~~~\wedge$ explored \green{e} $\cap$ set (stack \green{e}) = \{\}\\
    $~~~~\wedge$ ($\forall$ \green{v} \green{w}. \green{w} $\in$ $\mathcal{S}$ \green{e} \green{v} $\longleftrightarrow$ ($\mathcal{S}$ \green{e} \green{v} = $\mathcal{S}$ \green{e} \green{w}))\\
    $~~~~\wedge$ ($\forall$\green{v} $\in$ set (stack \green{e}).$\forall$ \green{w} $\in$ set (stack \green{e}).\green{v} $\neq$ \green{w} $\longrightarrow$ $\mathcal{S}$ \green{e} \green{v} $\cap$ $\mathcal{S}$ \green{e} \green{w} = \{\})\\
    $~~~~\wedge$ ($\forall$ \green{v}. \green{v} $\notin$ visited \green{e} $\longrightarrow$ $\mathcal{S}$ \green{e} \green{v} = \{\green{v}\})\\
    $~~~~\wedge$ $\bigcup$ \{$\mathcal{S}$ \green{e} \green{v} $|$ \green{v}. \green{v} $\in$ set (stack \green{e})\} = visited \green{e} - explored \green{e}\\
    $~~~~\wedge$ ($\forall$ \green{x} \green{y}. \green{x} $\preceq$ \green{y} in stack \green{e} $\longrightarrow$ reachable \green{y} \green{x})\\
    $~~~~\wedge$ ($\forall$ \green{x}. is\_subscc ($\mathcal{S}$ \green{e} \green{x}))\\
    $~~~~\wedge$ ($\forall$ \green{x} $\in$ explored \green{e}. $\forall$ \green{y}. reachable \green{x} \green{y} $\longrightarrow$ \green{y} $\in$ explored \green{e})\\
    $~~~~\wedge$ ($\forall$ \green{S} $\in$ sccs \green{e}. is\_scc \green{S})"
}

@@

\isabelle{
\bblue{definition} sub\_env \where\\
"\blue{sub\_env} \green{e e'} $\equiv$\\
$~~~~$visited \green{e} $\subseteq$ visited \green{e'}\\
$~~$$\wedge$ explored \green{e} $\subseteq$ explored \green{e'}\\
$~~$$\wedge$ ($\forall$ \green{v}. $\mathcal{S}$ \green{e v} $\subseteq$ $\mathcal{S}$ \green{e' v})\\
$~~$$\wedge$ ($\bigcup$\{$\mathcal{S}$ \green{e v} $|$ \green{v}. \green{v} $\in$ set (stack \green{e})\}) $\subseteq$ ($\bigcup$\{$\mathcal{S}$ \green{e' v} $|$ \green{v}. \green{v} $\in$ set (stack \green{e'})\})"
}

@@

\isabelle{
\bblue{definition} post\_dfs \where\\ 
"post\_dfs \green{v prev\_e e} $\equiv$\\
$~~~~$wf\_env \green{e}\\
$~~\wedge$ ($\forall$ \green{x}. reachable \green{v} \green{x} $\longrightarrow$ \green{x} $\in$ visited e)\\
$~~\wedge$ sub\_env \green{prev\_e} \green{e}\\
$~~\wedge$ ($\forall$ \green{n} $\in$ set (stack \green{e}). reachable \green{n} \green{v})\\
$~~\wedge$ ($\exists$ \green{ns}. stack \green{prev\_e} = \green{ns} @ (stack \green{e}))\\
$~~\wedge$ ($\forall$ \green{m n}. $\stackprec{\green{m}}{\green{n}}{\green{prec\_e}}$ $\longrightarrow$ \\ $~~~~~~$ ($\forall$ \green{u} $\in$ $\mathcal{S}$ \green{prev\_e m}. reachable \green{u} \green{v} $\wedge$ reachable \green{v} \green{n} $\longrightarrow$ $\mathcal{S}$ \green{e m} = $\mathcal{S}$ \green{e n}))\\
$~~\wedge$ ((\green{v} $\in$ explored \green{e} $\wedge$ stack \green{e} = stack \green{prev\_e}) $\lor$ \\
$~~~~~~$ (\green{v} $\in$ $\mathcal{S}$ \green{e} (hd (stack \green{e}))) $\wedge$ \\
$~~~~~~$ ($\exists$ \green{n} $\in$ set (stack \green{prev\_e}). $\mathcal{S}$ \green{e v} = $\mathcal{S}$ \green{e n}))"
}

@@

\isabelle{
\bblue{definition} pre\_dfss \where{}\\
"pre\_dfss \green{v vs e} $\equiv$\\
$~~~~$wf\_env \green{e}\\
$~~\land$ \green{v} $\in$ visited \green{e}\\
$~~\land$ \green{vs} $\subseteq$ successors \green{v}\\
$~~\land$ ($\forall$ \green{n} $\in$ set (stack \green{e}). reachable \green{n v})\\
$~~\land$ (stack \green{e} $\neq$ [])\\
$~~\land$ ($\forall$ \green{n}$\in$set(stack \green{e}).reachable \green{v n} $\longrightarrow$ \green{v}$\in$$\mathcal{S}$ \green{e n} $\lor$ ($\exists$ \green{m} $\in$ \green{vs}.reachable \green{m n}))\\
$~~\land$ (\green{v} $\in$ $\mathcal{S}$ \green{e} (hd (stack \green{e})))"
}

@@

\isabelle{
\bblue{definition} post\_dfss \where\\
"post\_dfss \green{v vs prev\_e e} $\equiv$\\
$~~~~$wf\_env \green{e}\\
$~~\land$ ($\forall$ \green{w} $\in$ \green{vs}. $\forall$ \green{x}. reachable \green{w x} $\longrightarrow$ \green{x} $\in$ visited \green{e})\\
$~~\land$ sub\_env \green{prev\_e e}\\
$~~\land$ ($\forall$ \green{n} $\in$ set (stack \green{e}). reachable \green{n v})\\
$~~\land$ (stack \green{e} $\neq$ [])\\
$~~\land$ ($\forall$ \green{n} $\in$ set (stack \green{e}). reachable \green{v n} $\longrightarrow$ \green{v} $\in$ $\mathcal{S}$ \green{e n})\\
$~~\land$ ($\exists$ \green{ns}. stack \green{prev\_e} = \green{ns} @ (stack \green{e}))\\
$~~\land$ (\green{v} $\in$ $\mathcal{S}$ \green{e} (hd (stack \green{e})))"
}
